\class : Week \psetnum

### 2022-10-31 541_ism/week5/solutions.tex
start ISM week 5

\newcommand{\psetnum}{0}
\newcommand{\class}{ASTR 541 - Interstellar Medium}

@@

\question{\textbf{1. Introduction to (Baryonic) Accounting}}

\noindent\emph{The general aim of this problem is to prove that the majority of baryonic matter does not lie in stars in galaxy or galaxies in clusters in a back-of-the-envelope calculation}
\answer{
    This answer heavily follows the solution laid out by \href{https://arxiv.org/pdf/astro-ph/9712020.pdf}{the paper that is the linked} in the question. To begin, we are given that $\Omega_B = 0.0455$ at $z = 0$. We can multiply this by the critical density,
    \begin{equation}
        \rho_c = \frac{3 H_0^2}{8 \pi G} \approx 9 \times 10^{-27} \unit{kg}{m^{-3}},
    \end{equation}
    to find that the mass density of baryonic matter is
    \begin{equation}
        \rho_B \approx 4 \times 10^{-28} \unit{kg}{m^{-3}}.
    \end{equation}

    \noindent Next we can calculate the mass density of stars and galaxies and compare it to this value. We can consider that most luminosity is coming from either a spheroid or a disk component and write that the density for each is
    \begin{equation}
        \rho = \mathcal{L}_B f_B \avg{M / L_B},
    \end{equation}
    where $\mathcal{L}_B$ is the mean luminosity density, $f_B$ is the fraction of light produced by this component and $\avg{M / L_B}$ is the mean mass-to-light ratio for this component.

    Let's assume that the only two components are spheroids and discs. The paper gives us that the fractions and mass-to-light ratios (in $\unit{M_{\odot}} / \unit{L_\odot}$) for each of these are $(0.385, 6.5)$ and $(0.582, 1.5)$ respectively. Summing these gives that the total mass density is
    \begin{equation}
        \rho = \mathcal{L}_B (0.385 \cdot 6.5 + 0.582 \cdot 1.5) \unit{M_{\odot}} / \unit{L_\odot} \approx 3.4 \mathcal{L}_B \unit{M_{\odot}} / \unit{L_\odot},
    \end{equation}
    From the paper we have that the mean luminosity density is approximately (assuming $h \approx 1$)
    \begin{equation}
        \mathcal{L}_B \approx 2 \times 10^{8} \unit{L_\odot}{Mpc^{-3}}
    \end{equation}
    Plugging this back in we get that the mass density of stars is
    \begin{equation}
        \rho_{\rm stars} \approx 6.8 \times 10^8 \unit{M_\odot}{Mpc^{-3}}
    \end{equation}\\

    \noindent We can now compare $\rho_B$ and $\rho_{\rm stars}$ and note that
    \begin{equation}
        \boxed{ \frac{\rho_{\rm stars}}{\rho_B} \approx 11\% }
    \end{equation}
    So from our back-of-the-envelope estimate we find that stars represent a small fraction of the baryonic matter within the Universe!!
}

>>>>

\newcommand{\psetnum}{5}
\newcommand{\class}{ASTR 541 - Interstellar Medium}

@@

\question{\textbf{1. }}
